### 2021-03-25 ch02_scientific_background/scientific_background.tex
Further improvements of text

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{
    ch02_scientific_background/figs/feedbacks_overview.pdf}
  \caption[
    Climate feedbacks and corresponding time scales.
  ]{
    Climate feedbacks and corresponding time scales. \enquote{+} refers to
    positive feedbacks, which amplify the effect of the external forcing (\eg{}
    the water vapor feedback). \enquote{-} refers to negative feedbacks, which
    diminish the effect of the external forcing (\eg{} the longwave radiation
    feedback). \enquote{+/-} refers to feedbacks which might be either positive
    or negative (\eg{} the cloud feedback). The smaller box highlights the
    large differences in time scales for the various feedbacks.
    \IPCCReproducedFrom{Cubasch2013}{their figure 1.2}.
  }
  \label{fig:02:feedbacks_overview}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch02_scientific_background/figs/keeling_curve.pdf}
  \caption[
    The Keeling Curve.
  ]{
    The Keeling Curve: monthly mean atmospheric \acs{CO2} concentration at the
    Mauna Loa Observatory, Hawaii ($19.5 \unit{\degree N}$, $155.6
    \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2019
    \autocite{Keeling2005}. The steady increase of the atmospheric \acs{CO2}
    concentration is superimposed with a pronounced seasonal oscillation caused
    by the seasonal \acs{CO2} cycle (see
    \cref{subsec:02:carbon_cycle_overview}).
  }
  \label{fig:02:keeling_curve}
\end{figure}

@@

In the last two decades, many emergent constraints on various aspects of the
Earth system have been published. Early studies tackled the hydrological cycle
\autocite{Allen2002} and the snow-albedo feedback \autocite{Hall2006}. Over the
years, the climate sensitivity expressed by the \ac{ECS} has been a prominent
target variable. Since cloud feedbacks are a major source of uncertainty for
climate sensitivity, a variety of papers focus on constraining \ac{ECS} with
cloud-related processes \autocite{Brient2015, Brient2016, Fasullo2012,
  Lipat2017, Sherwood2014, Su2014, Tian2015, Volodin2008, Qu2013, Zhai2015},
which are discussed in detail in
\cref{sec:05:comparison_of_emergent_constraints}. More recent studies aim to
constrain \ac{ECS} with the historical temperature variability
\autocite{Cox2018} or the historical warming trend
\autocite{JimenezdelaCuesta2019, Nijsse2020, Tokarska2020}. Emergent
constraints are not only limited to physical processes, but can also be applied
to other domains, like the global carbon cycle \autocite{Cox2013,
  Kwiatkowski2017, Wenzel2014, Wenzel2016, Winkler2019}. An extensive
discussion on the emergent constraint by \textcite{Wenzel2016}, which focuses
on the concentration-carbon feedback, is given in \cref{subsec:06:step_1}.

>>>>

\begin{figure}[!t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{
    ch02_scientific_background/figs/feedbacks_overview.pdf}
  \caption[
    Climate feedbacks and corresponding time scales.
  ]{
    Climate feedbacks and corresponding time scales. The different sings refer
    to the signs of the different feedbacks. Positive feedbacks amplify the
    effect of the external forcing (\eg{} the water vapor feedback) and
    negative feedbacks diminish the effect of the external forcing (\eg{} the
    longwave radiation feedback). An example for a feedback that might be
    either positive or negative is the cloud feedback. The smaller box
    highlights the large differences in time scales for the various feedbacks.
    \IPCCReproducedFrom{Cubasch2013}{their figure 1.2}.
  }
  \label{fig:02:feedbacks_overview}
\end{figure}

@@

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\EvenLargerFigureWidth{}]{
    ch02_scientific_background/figs/keeling_curve.pdf}
  \caption[
    The Keeling Curve.
  ]{
    The Keeling Curve: monthly mean atmospheric \acs{CO2} concentration at the
    Mauna Loa Observatory, Hawaii ($19.5 \unit{\degree N}$, $155.6
    \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2019
    \autocite{Keeling2005}. The steady increase of the atmospheric \acs{CO2}
    concentration is superimposed with a pronounced seasonal oscillation caused
    by the seasonal \acs{CO2} cycle (see
    \cref{subsec:02:carbon_cycle_overview}).
  }
  \label{fig:02:keeling_curve}
\end{figure}

@@

In the last two decades, many emergent constraints on various aspects of the
Earth system have been published. Early studies tackled the hydrological cycle
\autocite{Allen2002} and the snow-albedo feedback \autocite{Hall2006}. Over the
years, the climate sensitivity expressed by the \ac{ECS} has been a prominent
target variable. Since cloud feedbacks are a major source of uncertainty for
climate sensitivity, a variety of scientific papers focus on constraining
\ac{ECS} with cloud-related processes \autocite{Brient2015, Brient2016,
  Fasullo2012, Lipat2017, Sherwood2014, Su2014, Tian2015, Volodin2008, Qu2013,
  Zhai2015}, which are discussed in detail in
\cref{sec:05:comparison_of_emergent_constraints}. More recent studies aim to
constrain \ac{ECS} with the historical temperature variability
\autocite{Cox2018} or the historical warming trend
\autocite{JimenezdelaCuesta2019, Nijsse2020, Tokarska2020}. Emergent
constraints are not only limited to physical processes, but can also be applied
to other domains, like the global carbon cycle \autocite{Cox2013,
  Kwiatkowski2017, Wenzel2014, Wenzel2016, Winkler2019}. An extensive
discussion on the emergent constraint by \textcite{Wenzel2016}, which focuses
on the concentration-carbon feedback, is given in \cref{subsec:06:step_1}.

### 2021-03-25 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Further improvements of text

Various improvements of the underlying physical, biological and chemical
processes have been introduced to the \acs{CMIP}6 \acp{ESM} in order to
represent the coupled Earth system in more detail. Since many of these
processes directly influence the models' \ac{ECS} \autocite{Forster2020},
determining reasons for the increased \ac{ECS} in the \acs{CMIP}6 ensemble is
highly non-trivial. As discussed in detail by \textcite{Meehl2020}, one
possible reason is the consideration of aerosol-cloud interactions. New
prognostic aerosol schemes added to some \acs{CMIP}6 models that include
aerosol-cloud interactions could have produced overly large negative radiative
forcing, which then required a stronger model response to \ac{GHG} forcing in
order to reproduce the observed historical temperature trend. In fact,
\textcite{Meehl2020} find a weak inter-model relationship between the aerosol
forcing and \ac{ECS} for some \acs{CMIP}6 models, with larger negative
present-day aerosol forcing associated with higher \ac{ECS}. Such relationships
have also been found for previous climate model generations
\autocite{Kiehl2007, Forster2013}. However, due to the varying aerosol forcing
over time, these relations cannot be used to tune \ac{ECS} based on the models'
responses to aerosols over the different periods of the \nth{20} century
\autocite{Dittus2020}. A further type of aerosol-cloud interactions is for
example present in the \ac{HadGEM} version 3. The new aerosol scheme in this
model suppresses a strong negative cloud feedback over the Southern Ocean that
is found in the model's predecessor versions \autocite{BodasSalcedo2019}. The
Southern Ocean is one of the most pristine aerosol regions in the world
\autocite{Hamilton2014}, where aerosols predominantly originate from natural
sources like marine emissions of gaseous dimethyl sulfide. Earlier \ac{HadGEM}
versions show a large reduction in cloud droplet size with warming over the
Southern Ocean, which increases the cloud albedo and thus produces a negative
shortwave cloud feedback. On the contrary, in \acs{HadGEM}3, this negative
feedback is completely suppressed or even positive due to a small increase of
the cloud droplet size with warming, which eventually increases \ac{ECS} due to
an increased net cloud feedback \autocite{BodasSalcedo2019}.

@@

However, it is well known that the low-level mixed-phase clouds over the
Southern Ocean in these earlier climate models are biased towards too large
amounts of ice crystals and too little amounts of supercooled liquid water when
compared to satellite observations \autocite{BodasSalcedo2016}. Thus, the size
of the resulting strong negative cloud phase change feedback in these models
has long been questioned \autocite{McCoy2015, Tan2016}. Improvements in the
representation of mixed-phase clouds are known to reduce this long-standing
cloud phase bias over the Southern Ocean \autocite{BodasSalcedo2016, McCoy2016}
and to improve the representation of both cloud microphysical structure and
cloud radiative impacts in this region \autocite{Hyder2018, Kay2016}. For this
reason, the microphysical representation of mixed-phase clouds has been updated
in many \acs{CMIP}6 models. For example, in \acs{CESM}2, a new mixed-phase ice
nucleation scheme by \textcite{Hoose2010} replaces the \textcite{Meyers1992}
empirical scheme used in predecessor models \autocite{Gettelman2019}.
Simulations with the \textcite{Meyers1992} scheme show almost no supercooled
liquid water in high-latitude clouds and a strong negative cloud phase feedback
when ice turns to liquid. On the contrary, simulations with the new ice
nucleation scheme by \textcite{Hoose2010} show a strong reduction in this
feedback and are more consistent with observations of the current climate
\autocite{Gettelman2019}. Another example is the \acs{HadGEM}3 model, for which
a new cloud scheme that accounts for turbulent production of liquid water in
mixed-phase clouds has been implemented \autocite{BodasSalcedo2019}. The new
scheme parameterizes the role of subgrid-scale turbulence in the production and
maintenance of supercooled liquid water \autocite{Furtado2016}. As a
consequence, the cloud liquid water content over the Southern Ocean is
increased, which is in closer agreement to observations compared to earlier
\ac{HadGEM} versions. Moreover, the increase in the cloud liquid water content
with warming is reduced in \acs{HadGEM}3, leading to a reduction in the
negative cloud phase change feedback over the Southern Ocean
\autocite{BodasSalcedo2019}. By evaluating traditional metrics used for
numerical weather prediction, \textcite{Williams2020} additionally demonstrate
that short-term forecasts created with \acs{HadGEM}3 show a higher performance
than comparable simulations of predecessor model versions, which indicates an
improvement of the physical processes in the new climate model.

>>>>

Various improvements of the underlying physical, biological and chemical
processes have been introduced to the \acs{CMIP}6 \acp{ESM} in order to
represent the coupled Earth system in more detail. Since many of these
processes directly influence the models' \ac{ECS} \autocite{Forster2020},
determining reasons for the increased \ac{ECS} in the \acs{CMIP}6 ensemble is
highly non-trivial. As discussed in detail by \textcite{Meehl2020}, one
possible reason is the consideration of aerosol-cloud interactions. New
prognostic aerosol schemes added to some \acs{CMIP}6 models that include
aerosol-cloud interactions could have produced overly large negative radiative
forcing, which then required a stronger model response to \ac{GHG} forcing in
order to reproduce the observed historical temperature trend. In fact,
\textcite{Meehl2020} find a weak inter-model relationship between the aerosol
forcing and \ac{ECS} for some \acs{CMIP}6 models, with larger negative
present-day aerosol forcing associated with higher \ac{ECS}. Such relationships
have also been found for previous climate model generations
\autocite{Kiehl2007, Forster2013}. However, due to the varying aerosol forcing
over time, these relations cannot be used to tune \ac{ECS} based on the models'
responses to aerosols over the different periods of the \nth{20} century
\autocite{Dittus2020}. Further aerosol-cloud interactions that are relevant in
the context of high climate sensitivity in \acs{CMIP}6 are found for example in
the \ac{HadGEM} version 3. The new aerosol scheme in this model suppresses a
strong negative cloud feedback over the Southern Ocean that is found in the
model's predecessor versions \autocite{BodasSalcedo2019}. The Southern Ocean is
one of the most pristine aerosol regions in the world \autocite{Hamilton2014},
where aerosols predominantly originate from natural sources like marine
emissions of gaseous dimethyl sulfide. Earlier \ac{HadGEM} versions show a
large reduction in cloud droplet size with warming over the Southern Ocean,
which increases the cloud albedo and thus produces a negative shortwave cloud
feedback. On the contrary, in \acs{HadGEM}3, this negative feedback is
completely suppressed or even positive due to a small increase of the cloud
droplet size with warming, which eventually increases \ac{ECS} due to an
increased net cloud feedback \autocite{BodasSalcedo2019}.

@@

However, it is well known that the low-level mixed-phase clouds over the
Southern Ocean in these earlier climate models are biased towards too large
amounts of ice crystals and too little amounts of supercooled liquid water when
compared to satellite observations \autocite{BodasSalcedo2016}. Thus, the size
of the resulting strong negative cloud phase change feedback in these models
has long been questioned \autocite{McCoy2015, Tan2016}. Improvements in the
representation of mixed-phase clouds are known to reduce this long-standing
cloud phase bias over the Southern Ocean \autocite{BodasSalcedo2016, McCoy2016}
and to improve the representation of both cloud microphysical structure and
cloud radiative impacts in this region \autocite{Hyder2018, Kay2016}. For this
reason, the microphysical representation of mixed-phase clouds has been updated
in many \acs{CMIP}6 models. For example, in \acs{CESM}2, a new mixed-phase ice
nucleation scheme by \textcite{Hoose2010} replaces the \textcite{Meyers1992}
empirical scheme used in predecessor models \autocite{Gettelman2019}.
Simulations with the \textcite{Meyers1992} scheme show almost no supercooled
liquid water in high-latitude clouds and a strong negative cloud phase feedback
when ice turns to liquid. On the contrary, simulations with the new ice
nucleation scheme by \textcite{Hoose2010} show a strong reduction in this
feedback. Since the new scheme also reduces significant cloud phase biases over
the Southern Ocean that are present in the old scheme, the reduced negative
cloud phase feedback is thought to be more consistent with observations of the
current climate \autocite{Gettelman2019}. Another example is the \acs{HadGEM}3
model, for which a new cloud scheme that accounts for turbulent production of
liquid water in mixed-phase clouds has been implemented
\autocite{BodasSalcedo2019}. The new scheme parameterizes the role of
subgrid-scale turbulence in the production and maintenance of supercooled
liquid water \autocite{Furtado2016}. As a consequence, the cloud liquid water
path over the Southern Ocean is increased, which is more in line with
observations from microwave satellite retrievals. Moreover, the increase in the
cloud liquid water path with warming is reduced in \acs{HadGEM}3, leading to a
reduction in the negative cloud phase change feedback over the Southern Ocean
\autocite{BodasSalcedo2019}. In addition, \textcite{Williams2020} show that
short-term forecasts created with \acs{HadGEM}3 are more accurate than
comparable simulations from predecessor model versions (using traditional
metrics from numerical weather prediction), which indicates an improvement of
the physical processes in the new climate model.

### 2021-03-25 ch05_paper_ecs/paper_ecs.tex
Further improvements of text

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch05_paper_ecs/figs/constrained_ranges_summary.pdf}
  \caption[
    Overview of the constrained \acf{ECS} ranges given by the eleven analyzed
    emergent constraints using the \acs{CMIP}5 and \acs{CMIP}6 ensemble.
  ]{
    Overview of the constrained \acf{ECS} ranges given by the eleven analyzed
    emergent constraints using the \acs{CMIP}5 (blue) and \acs{CMIP}6 (orange)
    ensemble. The bars correspond to the $66 \unit{\%}$ confidence intervals
    ($\rangeunit{17}{83}{\%}$) calculated from the constrained \acl{PDF} of
    \acs{ECS} following \cref{eq:02:pdf_y_given_x0}. The vertical black lines
    illustrate the corresponding best estimates of the distributions. For
    almost all emergent constraints, the constrained best estimate and range of
    \ac{ECS} is higher in the \acs{CMIP}6 ensemble.
  }
  \label{fig:05:overview_results}
\end{figure}

@@

This paper assesses eleven different emergent constraints on \ac{ECS}, of which
most are directly or indirectly related to cloud feedbacks, by applying them to
results from \acp{ESM} contributing to \acs{CMIP}5 and \acs{CMIP}6. Of
particular interest are the results from \acs{CMIP}6, since all analyzed
emergent constraints have been published prior to the availability of
\acs{CMIP}6 data. In summary, the best estimate of \ac{ECS} averaged over all
emergent constraints increases by $12 \unit{\%}$ when moving from relationships
trained on \acs{CMIP}5 to those trained on \acs{CMIP}6. Some increase is
predicted by almost every constraint we analyze (with TII as the only
exception) and can be at least partly explained by the increased multi-model
mean \ac{ECS} of \acs{CMIP}6, which has not been accompanied by systematic
changes in the constraint variables that could explain this increase, leading
to regression fits with higher intercept values at observed constraint values.
This is also illustrated by the \acs{CMIP}5 and \acs{CMIP}6 \acp{MMM} in the
left columns of
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}
(colored circles), in which the connecting line between the \acs{CMIP}5 and
\acs{CMIP}6 \ac{MMM} is not parallel to the \acs{CMIP}5 emergent relationships
for all emergent constraints.

>>>>

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch05_paper_ecs/figs/constrained_ranges_summary.pdf}
  \caption[
    Overview of the constrained \acf{ECS} ranges given by the eleven emergent
    constraints using the \acs{CMIP}5 and \acs{CMIP}6 ensemble.
  ]{
    Overview of the constrained \acf{ECS} ranges given by the eleven analyzed
    emergent constraints using the \acs{CMIP}5 (blue) and \acs{CMIP}6 (orange)
    ensemble. The bars correspond to the $66 \unit{\%}$ confidence intervals
    ($\rangeunit{17}{83}{\%}$) calculated from the constrained \acl{PDF} of
    \acs{ECS} following \cref{eq:02:pdf_y_given_x0}. The vertical black lines
    illustrate the corresponding best estimates of the distributions. For
    almost all emergent constraints, the constrained best estimate and range of
    \ac{ECS} is higher in the \acs{CMIP}6 ensemble.
  }
  \label{fig:05:overview_results}
\end{figure}

@@

In this \namecref{ch:05:paper_ecs}, eleven different emergent constraints on
\ac{ECS}, which are mostly directly or indirectly related to cloud feedbacks,
are assessed by applying them to results from \acp{ESM} contributing to
\acs{CMIP}5 and \acs{CMIP}6. Of particular interest are the results from
\acs{CMIP}6, since all analyzed emergent constraints have been published prior
to the availability of \acs{CMIP}6 data. In summary, the best estimate of
\ac{ECS} averaged over all emergent constraints increases by $12 \unit{\%}$
when moving from relationships trained on \acs{CMIP}5 to those trained on
\acs{CMIP}6. Some increase is predicted by almost every constraint we analyze
(with TII as the only exception) and can be at least partly explained by the
increased multi-model mean \ac{ECS} of \acs{CMIP}6, which has not been
accompanied by systematic changes in the constraint variables that could
explain this increase, leading to regression fits with higher intercept values
at observed constraint values. This is also illustrated by the \acs{CMIP}5 and
\acs{CMIP}6 \acp{MMM} in the left columns of
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}
(colored circles), in which the connecting line between the \acs{CMIP}5 and
\acs{CMIP}6 \ac{MMM} is not parallel to the \acs{CMIP}5 emergent relationships
for all emergent constraints.

### 2021-03-25 ch06_paper_gpp/paper_gpp.tex
Further improvements of text

In this paper, we develop a two-step approach to constrain the projected
\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario. In the first step, we constrain the global mean \ac{GPP}
to $\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint
approach which we assume can also be applied to the global mean \ac{GPP} at the
end of the \nth{21} century in the \acs{RCP}8.5 scenario (see discussion in
\cref{subsec:06:step_1}). This step corresponds to the correction of the
\acp{ESM}' biases in the response of future \ac{GPP} to rising atmospheric
\ac{CO2} concentration, which is the main source of uncertainty in future
\ac{GPP} projections \autocite{Arora2013, Haverd2020, Rogers2017}. As already
discussed in \cref{subsec:06:step_1}, due to other than \ac{CO2} fertilization
drivers of the increase in the \ac{CO2} seasonal cycle amplitude
\autocite{Bastos2019, Forkel2016, Piao2018, Zhao2016} this emergent constraint
is not undisputed and can be replaced with another emergent constraint in
future studies. In the second step, a \ac{ML} approach is used to further
constrain the gridded \ac{GPP} based on process-based present-day predictors.
We consider two target variables: first (step 2a) the gridded monthly
climatologies of absolute \ac{GPP} (\range{2091}{2100}) and second (step 2b)
the gridded fractional \ac{GPP} change over the \nth{21} century (years
\range{2091}{2100} versus years \range{1991}{2000}; see \cref{eq:06:f}). Both
approaches give consistent results (see

>>>>

In this \namecref{ch:06:paper_gpp}, a two-step approach to constrain the
projected \ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in
the \acs{RCP}8.5 scenario is developed. In the first step, we constrain the
global mean \ac{GPP} to $\pmrangeunit{171}{12}{\GtCyr{}}$ using a published
emergent constraint approach which we assume can also be applied to the global
mean \ac{GPP} at the end of the \nth{21} century in the \acs{RCP}8.5 scenario
(see discussion in \cref{subsec:06:step_1}). This step corresponds to the
correction of the \acp{ESM}' biases in the response of future \ac{GPP} to
rising atmospheric \ac{CO2} concentration, which is the main source of
uncertainty in future \ac{GPP} projections \autocite{Arora2013, Haverd2020,
  Rogers2017}. As already discussed in \cref{subsec:06:step_1}, due to other
than \ac{CO2} fertilization drivers of the increase in the \ac{CO2} seasonal
cycle amplitude \autocite{Bastos2019, Forkel2016, Piao2018, Zhao2016} this
emergent constraint is not undisputed and can be replaced with another emergent
constraint in future studies. In the second step, a \ac{ML} approach is used to
further constrain the gridded \ac{GPP} based on process-based present-day
predictors. We consider two target variables: first (step 2a) the gridded
monthly climatologies of absolute \ac{GPP} (\range{2091}{2100}) and second
(step 2b) the gridded fractional \ac{GPP} change over the \nth{21} century
(years \range{2091}{2100} versus years \range{1991}{2000}; see \cref{eq:06:f}).
Both approaches give consistent results (see

### 2021-03-25 ch07_conclusion/conclusion.tex
Further improvements of text

The analysis of future climate projections from numerical climate model
simulations is of paramount importance to assess future climate change under
different forcing scenarios. Since this involves metrics of public interest
like the allowable fossil fuel emissions to meet particular warming targets,
for example the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement
\autocite{UNFCCC2015}, research in this field of science is not only relevant
to climate scientists but to policymakers and the whole human society. In the
light of the large spread of climate sensitivity in the most recent generation
of climate models from \acs{CMIP}6 \autocite{Meehl2020}, a careful statistical
evaluation and refinement of the output of multi-model climate projections is
as relevant as ever. This thesis quantifies the associated uncertainties,
presents the evaluation of established methods to reduce these uncertainties in
the new climate model ensemble and finally introduces a novel alternative
technique based on supervised \ac{ML}.

@@

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. Therefore, this analysis does not focus on
reducing uncertainties in climate sensitivity itself but rather on a selected
process that contributes to it: \ac{GPP}. \ac{GPP} is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
from the atmosphere. In the first step of the new two-step approach, the global
mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
\acs{RCP}8.5 scenario is constrained to $\pmrangeunit{171}{12}{\GtCyr{}}$ using
a published emergent constraint by \textcite{Wenzel2016}. This first step
corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2}, which shows a
range of $\rangeunit{156}{247}{\GtCyr{}}$ in the raw multi-model ensemble. In
the second step, a \ac{ML}-based climate model weighting approach is used to
further constrain the gridded \ac{GPP} based on present-day predictors that are
relevant for the simulation of \ac{GPP} in the \acp{ESM}. The \ac{ML} approach
is mathematically similar to the \ac{MDER} approach \autocite{Karpechko2013,
  Senftleben2020, Wenzel2016a}, but additionally considers multi-dimensional
(gridded) target variables and non-linear relationships between the predictors
and the target variable. A a relationship between process‐oriented predictors
and future projections of \ac{GPP} is established and then utilized to project
today's observed conditions into the future. The prediction phase of the new
method can be interpreted as an implicit performance weighting. However, due to
the complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not
possible to extract specific values for the individual weights. Two target
variables are considered: the gridded monthly climatologies of absolute
\ac{GPP} (\range{2091}{2100}) and the gridded fractional \ac{GPP} change over
the \nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. The new approach is validated by
comparing it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) is found when using the \ac{ML} approach. Moreover,
the evaluation of the global and local feature importance allows further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by the
new approach. For the second variable (fractional change in \ac{GPP}), \ac{T}
and \ac{LAI} are the dominant features. This study directly addresses
\cref{enum:01:question_3} (\enquote{\emph{\KeyScienceQuestionThree{}}}), which
can be confidently answered with \enquote{yes} based on the results found.

>>>>

The analysis of future climate projections from numerical climate model
simulations is of paramount importance to assess future climate change under
different forcing scenarios. Since this involves metrics of public interest
like the allowable fossil fuel emissions to meet particular warming targets,
for example the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement
\autocite{UNFCCC2015}, research in this field of science is not only relevant
to climate scientists but to policymakers and the whole human society. In the
light of the large spread of climate sensitivity in the most recent generation
of climate models from \acs{CMIP}6, that has been demonstrated in
\cref{ch:04:papers_ecs_tcr_assessment}, a careful statistical evaluation and
refinement of the output of multi-model climate projections is as relevant as
ever. This thesis quantifies the associated uncertainties, presents the
evaluation of established methods to reduce these uncertainties in the new
climate model ensemble and finally introduces a novel alternative technique
based on supervised \ac{ML}.

@@

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. Therefore, this analysis does not focus on
reducing uncertainties in climate sensitivity itself but rather on a selected
process that contributes to it: \ac{GPP}. \ac{GPP} is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
from the atmosphere. In the first step of the new two-step approach, the global
mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
\acs{RCP}8.5 scenario is constrained to $\pmrangeunit{171}{12}{\GtCyr{}}$ using
a published emergent constraint by \textcite{Wenzel2016}. In contrast to that,
the corresponding multi-model range of \ac{GPP} in the raw \acs{CMIP}5 ensemble
is $\rangeunit{156}{247}{\GtCyr{}}$. This first step corrects the \acs{CMIP}5
models' \ac{GPP} response to \ac{CO2}, which is primarily driven by the
\ac{CO2} fertilization. In the second step, a \ac{ML}-based climate model
weighting approach is used to further constrain the gridded \ac{GPP} based on
present-day predictors that are relevant for the simulation of \ac{GPP} in the
\acp{ESM}. The \ac{ML} approach is mathematically similar to the \ac{MDER}
approach \autocite{Karpechko2013, Senftleben2020, Wenzel2016a}, but
additionally considers multi-dimensional (gridded) target variables and
non-linear relationships between the predictors and the target variable. A a
relationship between process‐oriented predictors and future projections of
\ac{GPP} is established and then utilized to project today's observed
conditions into the future. The prediction phase of the new method can be
interpreted as an implicit performance weighting. However, due to the complex
structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible to
extract specific values for the individual weights. Two target variables are
considered: the gridded monthly climatologies of absolute \ac{GPP}
(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
\nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. The new approach is validated by
comparing it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) is found when using the \ac{ML} approach. Moreover,
the evaluation of the global and local feature importance allows further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by the
new approach. For the second variable (fractional change in \ac{GPP}), \ac{T}
and \ac{LAI} are the most important features. A corresponding feature map shows
that \ac{T} dominates in the tropics and \ac{LAI} dominates in the northern
high latitudes, which might be attributed to an additional greening trend in
the high latitudes that enhances the \ac{GPP} increase in this regions compared
to regions closer to the equator through an extension of the growing season.
This study directly addresses \cref{enum:01:question_3}
(\enquote{\emph{\KeyScienceQuestionThree{}}}), which can be confidently
answered with \enquote{yes} based on the results found.
